\documentclass[11pt]{article}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.2em} % Adjust '1em' to your preference

\usepackage{geometry}
\geometry{margin=1in}

\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet}  % DO NOT CHANGE THIS

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{hyperref}
\usepackage{natbib}
\bibliographystyle{plainnat}

\usepackage{xcolor}
\newcommand{\todo}[1]{{\color{orange}{TODO: #1}}}
\newcommand{\note}[1]{{\color{blue}{NOTE: #1}}}
\newcommand{\gur}[1]{{\color{teal}{Gur: #1}}}
\newcommand{\yarden}[1]{{\color{magenta}{Yarden: #1}}}


\title{The Effect of Age of Enrollment on the Probability of Graduating in Courses}
\author{Gur Keinan 213635899 \and Yarden Adi 212585848}
\date{October 30, 2024}


\begin{document}

\maketitle

\begin{abstract}
    Many adolescents worldwide have wondered when they should start their journey in higher education. The writers of this project themselves had decided long ago to join the 'Atuda' program, which means starting university at the relatively young age of eighteen. Thus, it is natural to wonder - does the student's age at the start of learning at the university affect the success of the student? An answer to this question can drastically change the landscape of the campuses worldwide and the grades of those studying there. In this causal inference project, we aim to explore precisely that. Specifically, we investigate the following causal question: \emph{What is the causal effect of being an adult student (over 21) on the probability of completing academic courses?} Throughout this project, we present and explore the data we use to tackle this question and determine its suitability for causal analysis tasks, formally present the methods we use to answer the research question, present and discuss the analysis results, and conclude the project. One can find all of the resources used in the project in this \href{https://github.com/GurKeinan/Causal-Inference-Project-Effect-of-Age-on-Graduating}{GitHub repository}.
\end{abstract}

\textbf{Note on AI Usage:} We did not use generative AI to write this report. We only used ChatGPT, Claude, and Grammarly for revisions and corrections of our own writing.

\section{Data Review and Preprocessing}

\gur{Opening sentence.}

\subsection{Data Review}

The data used in our project originated from research done in Portugal by the \href{https://www.ipportalegre.pt/pt/}{Polytechnic Institute of Portalegre} (\gur{Equivalent to college}), as an attempt to provide information to the tutoring team about the risk of students' dropout and failure \citep{data7110146}. Commonly, it is used to build machine learning models for predicting academic performance and dropout (see relevant \href{https://www.kaggle.com/datasets/ankanhore545/dropout-or-academic-success/data}{Kaggle competitions in this subject}).

The data contains information about students' pre-academic background, age, academic performance, social and economic status, and other relevant variables. It consists of 4424 records and 37 variables \todo{number of categorical and numerical}, including the enrollment age. The dataset includes a trinary outcome variable indicating whether a student dropped out, graduated, or is still enrolled in the course. 

The data was created by joining 3 primary data sources: 
\begin{enumerate}
    \item CNAES (National Competition for Access to Higher Education) - contains information about students' academic backgrounds, demographics, and course applications at the time of their enrollment in Portuguese higher education institutions.
    \item AMS (Academic Management System) - provides student records data, including demographic information, course enrollments, and academic performance throughout their studies.
    \item PORDATA (Contemporary Portugal Database) - provides macroeconomic data, including unemployment, inflation, and GDP figures.
\end{enumerate}

\gur{Division of the data to classes like in table in the paper}


\todo{Figure star of treatment and target distributions, discussion.}

\todo{Age limitation in Portugal. We can mention and cite the fact that Portugal is said to welcome education in all ages. This statement is commonly used regarding older people, but still.}

\subsection{Data Preprocessing}

\paragraph{Treatment and Target variables} Following our causal question, we introduce a new binary variable - "being over 21". This variable acts as the \emph{treatment variable} in our analysis. We omit the numerical age variable for the rest of the analysis. Additionally, as our question focuses on the probability of completing academic courses, we introduce another binary variable - "Graduated from the course" rather than the commonly used trinary variable - "graduate/  enrolled/ dropout". This variable is the \emph{target variable} in our analysis.

\paragraph{Post-treatment variables} The dataset contains multiple records of students' academic accomplishments during their first year of higher education. Although informative for machine learning models trying to predict academic dropout, these accomplishments were recorded after the treatment was determined and, therefore, cannot be safely used in the causal inference procedure. We removed such variables, 

\paragraph{Clustering categorical values} The dataset is of impressive complexity and detail, as evident by the categorical variables with over 30 unique values. To perform meaningful analysis, we manually cluster similar categories into one broader category to represent them instead of simply removing the less frequent value. For example, we merged the values 'Armed Forces Professions', 'Armed Forces Officers', and 'Armed Forces Sergeants' into a single 'Armed Forces' value. We performed this procedure on five variables within the data set - the qualifications and occupations of the parents (mother and father) and the student's previous qualifications. 

\paragraph{Pruning categorical outliers} Even after performing the previous step, some variables still contained rare values. After careful consideration and visual inspections, we decided to prune some of the rare values to ease the analysis. Among the rest, the pruning resulted in only subjects with Portuguese nationality, \gur{detail here more}.


% Preprocessing:
% \begin{itemize}
%     \item Removing post-treatment variables with explanation.
%     \item Removing unnecessary variables for easier analysis.
%     \item Scaling variables? 
%     \item One-hot vectors for categorical variables?
%     \item Sub categorizing some variables for easier analysis. 
% \end{itemize}

\section{Assumptions for Causal Inference}

% \begin{itemize}
%     \item Formally present each condition.
%     \item Argue that it holds.
%     \item Common Support - using propensity scores.
%     \item Ignorability - using literature that presents facts and experiments that back up our assessment that we have the most relevant features.
%     \item SUTVA.
%     \item Consistency.
% \end{itemize}

In this section, we formally present and discuss four assumptions regarding the nature of our data. Combined, those assumptions guarantee the trustworthiness of an observational causal experiment's results.

\subsection{Stable Unit Treatment Value Assumption (SUTVA)}

The SUTVA assumption states that the potential outcomes of each unit are not affected by the treatment assignment of any other unit. This assumption is violated when there is interference between units, i.e., when the treatment of one unit affects the potential outcomes of other units. For example, if a treatment is a vaccine that prevents the spread of a disease, then the treatment of one individual can affect the potential outcomes of other individuals, as one individual's treatment can prevent the disease from spreading to others. In this case, the SUTVA assumption is violated.

\todo{In our data...}

\subsection{Consistency}

The consistency assumption states that an individual's potential outcome under his or her observed exposure history is the outcome that will actually be observed for that person. Put formally, for a unit that receives treatment $T$, we observe the corresponding potential outcome $Y = TY_1 + (1-T)Y_0$. This assumption can be violated when the individual's history is misreported \gur{(e.g., "Have you ever tried LSD?)}.

Our data originated from credible and regulated sources using careful and documented manipulations. Thus, we argue that the consistency assumption holds in our data.

\subsection{Ignorability – no unmeasured confounders}

The ignorability assumption states that the treatment assignment $T$ is independent of the potential outcomes $Y_0, Y_1$ given the observed covariates $X$, i.e., $Y_0, Y_1 \perp T \ | \ X$. 

This assumption can always be violated since it is impossible to effectively record or scrape all relevant data for the experiment. \gur{Literature!}

\subsection{Common support (overlap)}

The common support assumption states that each unit has a non-zero probability of receiving each treatment level, i.e., $\forall x \in X, P(T=1|X=x) > 0$ and $P(T=0|X=x) > 0$.

We empirically validate this assumption using propensity scores.

\section{Causal Analysis Methodology}

\begin{itemize}
    \item Present the relevant measures (ATE, etc).
    \item Present the methods we use for calculating the measures.
    \item Implementation details (a lot of them).
    \item Bootstrap.
\end{itemize}

\section{Results}

\begin{itemize}
    \item One graph to compare all of the results from the methods (box plots).
    \item Discussion.
    \item Analyze correlation / dominant features.
\end{itemize}

\section{Discussion}

\begin{itemize}
    \item Conclusions
    \item Possible weaknesses.
    \item Avenues for future work.
\end{itemize}

\bibliography{references.bib}

\appendix

\end{document}